% Created 2017-04-02 Sun 01:34
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\author{Compro Prasad}
\date{\today}
\title{}
\hypersetup{
 pdfauthor={Compro Prasad},
 pdftitle={},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.5)},
 pdflang={English}}
\begin{document}

\section*{Name}
\label{sec:orgbbd0645}
Abhishek Prasad
\section*{Email}
\label{sec:orga5e323a}
comproprasad@gmail.com
\section*{Name of the project}
\label{sec:org2a5ee67}
GNU Clisp
\section*{Summary}
\label{sec:org2ea2e3a}
Implementing thread safe hash tables in the core of GNU Clisp.
\section*{Benefits}
\label{sec:org9b30dde}
The community:
Having threading support is necessary everywhere from normal computing tasks to
larger applications running on production environments like servers. The problem
arises when the data being used in threads are not atomic in nature. Most of
data structures are thread safe in GNU Clisp except the hash tables which are
also extensively used in \href{https://sourceforge.net/p/clocc/hg/ci/default/tree/}{CLOCC}.

Me:
Since I am doing this project I will get to know how threads, mutexes and hash
tables are used in writing a program like GNU Clisp. And overall I would be
communicating with great people in the community I would learn other side stuff
too which will be an additional gift for me.
\section*{Things to be delivered}
\label{sec:orga02c250}
Since, my work consists of threading and hash table support so the files to be
dealt with are \texttt{src/zthread.d}, \texttt{src/xthread.d} and \texttt{src/hashtabl.d} and may be
other lisp files.

Preventing segfaults in MT builds while improving the hash table support. If
possible reimplementation of hash table will be done using Leapfrog probing
method so they will probably be implemented.
\section*{Plan}
\label{sec:org21094fd}
\begin{center}
\begin{tabular}{ll}
Date & Work\\
\hline
04-May to 30-May & Talking to the developer team, fixing small bugs and get acquainted with the project\\
 & \\
31-May to 11-June & Writing the lisp code to make simple threads and find out where the hash table hits a corner case. Get ready with enough tests.\\
 & \\
09-June to 30-June & Write functions to handle the thread safety in the hash tables.\\
 & \\
25-June to 15-July & Have the functions run against the test cases\\
 & \\
10-July to 20-July & Write the tests that were missed beforehand\\
 & \\
18-July to 25-July & Make the code more portable to run on different platforms\\
 & \\
23-July to 30-July & Polishing the code to be more readable for future development\\
\end{tabular}
\end{center}
\section*{Communication}
\label{sec:org5253a2c}
\href{https://sourceforge.net/p/clisp/mailman/clisp-devel/}{clisp-devel} mailing list archives on source forge is subscribed both via email
and RSS feeds. \href{https://sourceforge.net/p/clisp/mailman/clisp-announce/}{clisp-announce} is the place where announcements are made.
User bugs are available on \href{https://sourceforge.net/p/clisp/mailman/clisp-list/}{clisp-list}.

I have also asked some questions to the GNU Clisp developers who have helped me
most of the time but I am loving the new things that I learn when I communicate.
\section*{Qualification}
\label{sec:org5fc5c1d}
I have written a mathematical expression calculator, convex hull problem, AVL
trees and simple matrix operations. So, I am well acquainted with programming in
C/C++ using Makefiles. I also like designing/programming core stuff like a
compiler or a kernel or a DBMS.

Since, I am an Emacs user for around a year so I have started to like ELisp and
gradually lisp too.

\begin{itemize}
\item National Institute of Technology, Durgapur
\item Bachelor's course in Information Technology
\item Second(2nd year or 4th semester) year
\item Multi threading concept is in 4th semester
\item Hash Tables were taught in the 3rd semester
\item Basics of \texttt{pthread.h}
\end{itemize}
\end{document}
